---
title: "A prototype to process BoxCar data with MSnbase"
output:
    BiocStyle::html_document:
        toc_float: true
vignette: >
    %\VignetteIndexEntry{A prototype to process BoxCar data with MSnbase}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
    %\VignettePackage{MSnbaseBoxCar}
    %\VignetteDepends{MSnbase,BiocStyle,ggplot2}
---


```{r style, echo = FALSE, results = 'asis', message=FALSE}
BiocStyle::markdown()
```

# Preparation

Load required packages and functions. 

```{r pkgs, message = FALSE}
library("MSnbase")
library("MSnbaseBoxCar")
library("ggplot2")
```

Read a small dataset composed of 16 MS1 spectra as an `MSnExp`:

```{r data}
f <- dir(system.file("extdata", package = "MSnbaseBoxCar"),
         pattern = "boxcar.mzML",
         full.names = TRUE)
basename(f)
x <- readMSData(f, mode = "onDisk")
x
```

# Define boxcar groups

Define boxcar groups based on the `filterString` metadata variable:
full scans are encoded as `"FTMS + p NSI Full ms
[375.0000-1800.0000]"` while their respective Boxcar scans reflect the
provide adjacent m/z segments `"FTMS + p NSI SIM msx ms
[299.0000-1701.0000, 299.0000-351.0000, ...]"`


```{r filterStrings}
fData(x)$filterString[1:4]
```

The `bc_groups` function identifies full (noted `NA`) and BoxCar
spectra and groups the latter:


```{r bc_groups}
fData(x)$groups <- bc_groups(x)
fData(x)$groups
```
# Keep only BoxCar spectra

The next filter BoxCar spectra, as defined above.

```{r bc_is_boxcar}
xbc <- x[bc_is_boxcar(x)]
fData(xbc)$groups
```

```{r bc_plot1, fig.cap = "Beginning of the first adjacent BoxCar segments."}
bc_plot(xbc[1:3]) +
    xlim(440, 510)
```

# Combine BoxCar spectra 

Remove any peaks outside of the BoxCar segments.

```{r bc_zero_out_box}
xbc <- bc_zero_out_box(xbc, offset = 0.5)
xbc
```

```{r bc_plot2, fig.cap = "Peaks outside of the BoxCar segments have been removed."}
bc_plot(xbc[1:3]) +
    xlim(440, 510)
```

Combine BoxCar spectra to reconstitute the full scan and coerce result
back to an `MSnExp` object containing 4 spectra.


```{r combineSpectra}
res <- combineSpectra(xbc,
                      fcol = "groups",
                      method = boxcarCombine)
res
```

The above steps can also be piped into a single command.

```{r pipe, eval=FALSE}
library("magrittr")
res <- xbc %>%
    bc_zero_out_box(offset = 0.5) %>%
    combineSpectra(fcol = "groups",
                   method = boxcarCombine)

res
```

```{r plot, fig.cap = "Reconstructed full spectrum."}
plot(res[[1]])
```

The processed data can also be written to a new mzML file.

```{r write, eval=FALSE}
writeMSData(res, "boxcar_processed.mzML")
```

# References

- [BoxCar acquisition method enables single-shot proteomics at a depth
  of 10,000 proteins in 100
  minutes](https://www.nature.com/articles/s41592-018-0003-5).
- [`MSnbase` - an R/Bioconductor package for isobaric tagged mass
  spectrometry data visualization, processing and
  quantitation](https://pubmed.ncbi.nlm.nih.gov/22113085/).
- `MSnbase` [webpage](http://lgatto.github.io/MSnbase/).

# Session information

```{r si}
sessionInfo()
```
